---
title: 'Reproducible Research: Peer assessment 1'
author: "Clémence Mathelin"
date: "11 mai 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Repoducible Research - Peer Assessement 1
==========================================
This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October
and November, 2012 and include the number of steps taken in 5 minute intervals each day.

### The data
The variables included in this dataset are:

- steps: Number of steps taking in a 5-minute interval (missing values are coded as NA);
- date: The date on which the measurement was taken in YYYY-MM-DD format;
- interval: Identifier for the 5-minute interval in which measurement was taken.
The dataset is stored in a comma-separated-value (CSV) file and there are a total of 17,568 observations in this dataset.




## Loading and Processing the data
This following code reads the data into R and transform the data into a format suitable for the analysis (use factors and adapt the date format).

```{r load and process data}
unzip("activity.zip")
activity <- read.csv("./activity.csv")
activity$interval <- as.factor(activity$interval)
activity$date <- as.Date(activity$date)
```


## What is the mean total number of steps taken per day?
This part calculates the total number of steps taken per day and turns it into a histogram. Then, it calculates the mean and the median of the total number of steps tken per day.

1 Calculate the total number of steps taken per day
```{r total steps per day}
activityomit <- as.data.frame(activity[complete.cases(activity), ])
totalSteps <- tapply(activityomit$steps, activityomit$date, sum)
totalSteps
```

2 Make an histogram of the total number of steps taken each day
``` {r histogram per day}
hist(totalSteps,nrow(totalSteps), breaks = 10, main = "Frequency of Total Steps per Day")
```

3 Calculate and report the mean and median of the total number of steps taken per day
```{r mean and median}
meanActivity <- mean(totalSteps)
medianActivity <- median(totalSteps)
```

The **mean** of the total number of steps taken per day is `r meanActivity`.
The **median** of the total number of steps taken per day is `r medianActivity`.



## What is the average daily pattern?
The following code plots the average number of steps in each interval across all days, as well as the "busiest" (max) interval.

1 Make a time series plot of the 5-minute interval ans the average number of steps taken, averaged accross all days.
```{r time series plot}
dailyPattern <- aggregate(activityomit$steps, by = list(activityomit$interval), FUN = mean)
plot(dailyPattern, type = "l", 
     xlab = "5 min Intervals in a day", 
     ylab = "Average Number of Steps", 
     main = "The Average Daily Activity Pattern")
```

2 Which 5-minute interval, on average accross all the days in the dataset, contains the maximum number of steps?
```{r max interval}
maxInverval <- dailyPattern[which.max(dailyPattern[, 2]), 1]
```

The 5-minute interval that contains the maximum number of steps is the interval `r maxInterval`.



## Imputing missing values
This part has been created to underline possible bias into some calculations and summaries intoduced by the presence of missing values.

1 Calculate and report the total number of missing values in the data set.
```{r count NAs}
countNA <- sum(!complete.cases(activity))
```

There are `r countNA`rows with missing values in this dataset.


2 Devise a strategy for filling in all of the missing values in the dataset. 

The strategy we chose to impute missing values is to use the mean for the 5-minute interval.
To do so, we create a new dataset equal to the original data. However, the missing values are replaced by the corresponding 5-minute interval mean (see strategy above).


3 Create a new data set that is equal to the original dataset but with the missing values filled.
```{r new strategy}
newActivity <- activity
len1 <- nrow(newActivity)
len2 <- nrow(dailyPattern)
for (i in 1:len1) {
  if (is.na(newActivity$steps[i])) {
    for (j in 1:len2) {
      if (newActivity$interval[i] == dailyPattern[j, 1]) {
        newActivity$steps[i] = dailyPattern[j, 2]
      }
    } 
  }    
}
head(newActivity)
```


4 Make a histogram of the total number of steps taken each day and calculate and report the mean and median total number of steps taken per day.
```{r plot new strategy}
totalNewSteps <- tapply(newActivity$steps, newActivity$date, sum)
hist(totalNewSteps,nrow(totalNewSteps), breaks = 10, main = "Frequency of Total Steps per Day")
meanNewSteps <- mean(totalNewSteps, na.rm = TRUE)
medianNewSteps <- median(totalNewSteps, na.rm = TRUE)
```

With our new strategy, the **mean** of the total number of steps taken per day is `r meanNewSteps`, while our last mesurement was `r meanActivity`.
With our new strategy, the **median** of the total number of steps taken per day is `r medianNewSteps`.

